---
title: "User-led grants"
output: html_notebook
---

# Import data

Import the data corresponding to a search of 'user-led' in the Grant Nav website  (http://grantnav.threesixtygiving.org).

```{r}
userledgrantsfull <- read.csv("grantnav-userled-8469.csv", na.strings=c(""," ","NA","-","n/a"))
```

# Data cleaning

Clean column names

```{r}
colnames(userledgrantsfull)[1] <- "Identifier"
```

Select only the columns useful for this visualisation

```{r}
library(dplyr)
userledgrants <- select(userledgrantsfull, c("Identifier","Title","Description","Amount.Awarded","Award.Date","URL","Recipient.Org.Identifier","Recipient.Org.Name","Recipient.Org.Name","Recipient.Org.Charity.Number","Recipient.Org.Company.Number","Recipient.Org.Location.0.Geographic.Code.Type","Funding.Org.Identifier","Funding.Org.Name","Grant.Programme.Code","Grant.Programme.Title"))
```

Filter the rows related to user-led grants and user-led organizations. 

It's important that we add all the details because filtering only by the word "led" we would count in grants containing unrelated topics, such as "led lighting" or "led by X professor". 

On the other side, if we only filter by "user-led" we miss a lot of grants that fit in the description.

```{r}
userledgrantsdescr <- filter(userledgrants, grepl("user(  *|- *)led|community(- *| *)led|patient(- *|  *)led|youth(- *|  *)led|people(- *|  *)led|volunteer(- *|  *)led|locally(- *|  *)led|child(- *|  *)led|parent(- *|  *)led|peer(- *|  *)led|girl(- *|  *)led|wom(a|e)n(- *|  *)led|disabled(- *|  *)led|deaf(- *|  *)led|citizen(- *|  *)led|B(A|)ME(R|)(- *|  *)led|student(- *|  *)led|ethnic(- *|  *)led|resident(- *|  *)led|pupil(- *|  *)led|young people(- *|  *)led|member(- *|  *)led|offender(- *|  *)led|family(- *|  *)led|¡young person(- *|  *)led|survivor(- *|  *)led|seeker(- *|  *)led|led by (the |)community|led by (the |)patient|led by (the |)user|led by (the |)youth|led by (the |)people|led by (the |)volunteer|led by (the |)children|led by (the |)parent|led by (the |)peer|led by (the |)girl|led by (the |)women|led by (the |)deaf|led by (the |)disabled|led by (the |)BAME|led by (the |)student|led by (the |)resident|led by (the |)pupil|led by (the |)member|led by (the |)local|led by (the |a |)young|led by (the |)old|led by (the |)elder", userledgrants$Description, ignore.case = TRUE, perl = TRUE))

userledgrantstitle <- filter(userledgrants, grepl("user(  *|- *)led|community(- *| *)led|patient(- *|  *)led|youth(- *|  *)led|people(- *|  *)led|volunteer(- *|  *)led|locally(- *|  *)led|child(- *|  *)led|parent(- *|  *)led|peer(- *|  *)led|girl(- *|  *)led|wom(a|e)n(- *|  *)led|disabled(- *|  *)led|deaf(- *|  *)led|citizen(- *|  *)led|B(A|)ME(R|)(- *|  *)led|student(- *|  *)led|ethnic(- *|  *)led|resident(- *|  *)led|pupil(- *|  *)led|young people(- *|  *)led|member(- *|  *)led|offender(- *|  *)led|family(- *|  *)led|¡young person(- *|  *)led|survivor(- *|  *)led|seeker(- *|  *)led|led by (the |)community|led by (the |)patient|led by (the |)user|led by (the |)youth|led by (the |)people|led by (the |)volunteer|led by (the |)children|led by (the |)parent|led by (the |)peer|led by (the |)girl|led by (the |)women|led by (the |)deaf|led by (the |)disabled|led by (the |)BAME|led by (the |)student|led by (the |)resident|led by (the |)pupil|led by (the |)member|led by (the |)local|led by (the |a |)young|led by (the |)old|led by (the |)elder", userledgrants$Title, ignore.case = TRUE, perl = TRUE))
```

Merge rows that contain user-led related information in the description and those that contain it in the title.

```{r}
library(dplyr)
onlyuserledgrants <- union(userledgrantsdescr, userledgrantstitle)
```

Extract year when the grant was awarded

```{r}
onlyuserledgrants$Year <- substr(onlyuserledgrants$Award.Date, 1, 4)
onlyuserledgrants$Year <- as.numeric(onlyuserledgrants$Year)
```


# Categorising 

## By theme

Go through the grant description to add a category label to each grant. 

```{r}
childyoung <- ifelse(grepl("children|pupil|student|child|young|youth", onlyuserledgrants$Description, ignore.case = TRUE), 1, 0)
old <- ifelse(grepl(" old|elder", onlyuserledgrants$Description, ignore.case = TRUE), 1, 0)
community <- ifelse(grepl("community", onlyuserledgrants$Description, ignore.case = TRUE), 1, 0)
house <- ifelse(grepl("housing|homeless", onlyuserledgrants$Description, ignore.case = TRUE), 1, 0)
mentalhealth <- ifelse(grepl("mental health|phsycolog", onlyuserledgrants$Description, ignore.case = TRUE), 1, 0)
health <- ifelse(grepl("cardiac|cancer|stroke", onlyuserledgrants$Description, ignore.case = TRUE), 1, 0)
race <- ifelse(grepl(" race| B(A|)ME| racial| ethnic| minority", onlyuserledgrants$Description, ignore.case = TRUE), 1, 0)
women <- ifelse(grepl(" woman| girl| women|domestic abuse", onlyuserledgrants$Description, ignore.case = TRUE), 1, 0)
disabled <- ifelse(grepl("disability| deaf| disabled| disabilities", onlyuserledgrants$Description, ignore.case = TRUE), 1, 0)
themes <- data.frame(disabled,health,mentalhealth,race,house,women,old,childyoung,community,onlyuserledgrants$Description,onlyuserledgrants$Title)
themes <- mutate(themes, sum = childyoung + old + community + house + mentalhealth + health + race + women + disabled)
themes$theme <- colnames(themes[1:9])[apply(themes[1:9],1,which.max)]
themes$theme <- ifelse(themes$sum==0, "Not classified", themes$theme)
```

```{r}
filter(themes, sum==0)$theme
```

```{r}
sum(childyoung)
sum(old)
sum(community)
sum(house)
sum(mentalhealth)
sum(health)
sum(race)
sum(women)
sum(disabled)
```

```{r}
sort(table(themes$theme), decreasing=TRUE)
```

```{r}
onlyuserledgrants$Theme <- themes$theme
```

Add frequency of each theme. 

```{r}
frequencythemes <- plyr::count(themes, 'theme')
themesfreq <- dplyr::full_join(themes, frequencythemes, by.x = theme)
onlyuserledgrants$ThemeFreq <- themesfreq$freq
```

Substitute names for the ones to display

```{r}
onlyuserledgrants$Theme <- gsub("childyoung", "Children and youngs", onlyuserledgrants$Theme)
onlyuserledgrants$Theme <- gsub("old", "The elderly", onlyuserledgrants$Theme)
onlyuserledgrants$Theme <- gsub("community", "Community", onlyuserledgrants$Theme)
onlyuserledgrants$Theme <- gsub("house", "Housing", onlyuserledgrants$Theme)
onlyuserledgrants$Theme <- gsub("mentalhealth", "Mental health", onlyuserledgrants$Theme)
onlyuserledgrants$Theme <- gsub("health", "Health", onlyuserledgrants$Theme)
onlyuserledgrants$Theme <- gsub("race", "Black, Asian and minority ethnic", onlyuserledgrants$Theme)
onlyuserledgrants$Theme <- gsub("women", "Women", onlyuserledgrants$Theme)
onlyuserledgrants$Theme <- gsub("disabled", "People with disabilities", onlyuserledgrants$Theme)
```


## By organisation type

Matching each recipient organization to a type

```{r}
#If it has a charity number, it is a charity
charity <- ifelse(is.na(onlyuserledgrants$Recipient.Org.Charity.Number), 0, 1)

#For the rest, we look at the name of the organization
cic <- ifelse(grepl("CIC|C.I.C|Community Interest Company", onlyuserledgrants$Recipient.Org.Name, ignore.case = TRUE), 1, 0)
university <- ifelse(grepl("university|college", onlyuserledgrants$Recipient.Org.Name, ignore.case = TRUE), 1, 0)
la <- ifelse(grepl("county|council|borough", onlyuserledgrants$Recipient.Org.Name, ignore.case = TRUE), 1, 0)
commgroupasso <- ifelse(grepl("Community Group|Association", onlyuserledgrants$Recipient.Org.Name, ignore.case = TRUE), 1, 0)
school <- ifelse(grepl("School", onlyuserledgrants$Recipient.Org.Name, ignore.case = TRUE), 1, 0)
trust <- ifelse(grepl("Trust", onlyuserledgrants$Recipient.Org.Name, ignore.case = TRUE), 1, 0)
privatecompany <- ifelse(grepl("Ltd|Limited", onlyuserledgrants$Recipient.Org.Name, ignore.case = TRUE), 1, ifelse(!is.na(onlyuserledgrants$Recipient.Org.Company.Number), 1, 0))

orgtype <- data.frame(charity,cic,university,la,commgroupasso,school,trust,privatecompany)
orgtype <- mutate(orgtype, sum = charity + cic + university + la + commgroupasso + school + trust + privatecompany)
orgtype$orgtype <- colnames(orgtype[1:8])[apply(orgtype[1:8],1,which.max)]
orgtype$orgtype <- ifelse(orgtype$sum==0, "Not classified", orgtype$orgtype)
```

I anyone missing?
```{r}
filter(orgtype, charity == "No" & cic == "No" & university == "No" & la == "No" & commgroupasso == "No" & school == "No" & institute == "No" & trust == "No" & privatecompany == "No")
```

```{r}
sort(table(orgtype$orgtype), decreasing=TRUE)
```

```{r}
onlyuserledgrants$OrgType <- orgtype$orgtype
```

Add frequency of each organization
```{r}
frequencyorgtype <- plyr::count(orgtype, 'orgtype')
orgtypefreq <- dplyr::full_join(orgtype, frequencyorgtype, by.x = orgtype)
onlyuserledgrants$OrgTypeFreq <- orgtypefreq$freq
```

```{r}
frequencyorg <- plyr::count(onlyuserledgrants, 'Funding.Org.Name')
onlyuserledgrants <- dplyr::full_join(onlyuserledgrants, frequencyorg, by.x = Funding.Org.Name)
```

```{r}
colnames(onlyuserledgrants)[21] <- "OrgFreq"
```

Change to labels to display
```{r}
onlyuserledgrants$OrgType <- gsub("charity", "Charity", onlyuserledgrants$OrgType)
onlyuserledgrants$OrgType <- gsub("privatecompany", "Private company", onlyuserledgrants$OrgType)
onlyuserledgrants$OrgType <- gsub("school", "School", onlyuserledgrants$OrgType)
onlyuserledgrants$OrgType <- gsub("la", "Local authority", onlyuserledgrants$OrgType)
onlyuserledgrants$OrgType <- gsub("cic", "Community Interest Company", onlyuserledgrants$OrgType)
onlyuserledgrants$OrgType <- gsub("university", "Univerity or college", onlyuserledgrants$OrgType)
onlyuserledgrants$OrgType <- gsub("commgroupasso", "Community group or association", onlyuserledgrants$OrgType)
onlyuserledgrants$OrgType <- gsub("trust", "Trust", onlyuserledgrants$OrgType)
```

```{r}
onlyuserledgrants$OrgType <- gsub("Not cLocal authorityssified", "Not classified", onlyuserledgrants$OrgType)
```

##Amount per founder

Add a column with the amount each founder has awarded
```{r}
library(dplyr)
totalamount <-  onlyuserledgrants %>% group_by(FundingOrgName) %>% summarise(TotalAmountFounder = sum(AmountAwarded))
onlyuserledgrants <- dplyr::full_join(onlyuserledgrants, totalamount, by = "FundingOrgName")
```

# Export

Make sure we don't export blank rows
```{r}
onlyuserledgrants <- filter(onlyuserledgrants, !is.na(onlyuserledgrants$Identifier))
```

Clean column names
```{r}
colnames(onlyuserledgrants) <- gsub("[.]", "", colnames(onlyuserledgrants))
```

Export to a JSON file to visualise with D3.js
```{r}
library(jsonlite)
userledjson <- toJSON(onlyuserledgrants)
write(userledjson, "userledfinal.json")
```


# Analysis

```{r}
install.packages("summarytools")
```

```{r}
library(summarytools)
view(dfSummary(onlyuserledgrants))
```

Count user led-grants by year
```{r}
table(onlyuserledgrants$Year)
```

Count all the grants by year
```{r}
userledgrants$Year <- substr(userledgrants$Award.Date, 1, 4)
userledgrants$Year <- as.numeric(userledgrants$Year)
table(userledgrants$Year)
```

Count by theme
```{r}
sort(table(onlyuserledgrants$Theme), decreasing = T)
```

Percentage by theme
```{r}
sort(100*prop.table(table(onlyuserledgrants$Theme)), decreasing = T)
```

Total amount
```{r}
sum(onlyuserledgrants$AmountAwarded)
```


Max amount
```{r}
max(onlyuserledgrants$AmountAwarded)
dplyr::slice(onlyuserledgrants, which.max(onlyuserledgrants$AmountAwarded))
```

Median amount
```{r}
stats::median(onlyuserledgrants$AmountAwarded)
```

Count by founder
```{r}
sort(table(onlyuserledgrants$FundingOrgName), decreasing = T)
```

Percentage by founder
```{r}
sort(prop.table(table(onlyuserledgrants$FundingOrgName))*100, decreasing = T)
```

Amount by founder

```{r}
library(dplyr)
onlyuserledgrants %>% 
  group_by(FundingOrgName) %>% 
  summarise(TotalAmount = sum(AmountAwarded))
```

